---
title: "Visualizing Distributions: Lab 2"
author: "Tina Piltner"
date: "2025-06-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
Define variables containing the heights of males and females like this: 

```{r}
library(dslabs)
male <- heights$height[heights$sex == "Male"]
female <- heights$height[heights$sex == "Female"]
```

How many measurements do we have for each? 

```{r}
length(male)
length (female) 
```
812 measurements for male. 
238 measurements for female. 

## Question 2
Create a five row table showing female_percentiles and male_percentiles with the 10th, 30th, 50th, 70th, & 90th percentiles for each sex. Then create a data frame with these two as columns.

```{r}
percentiles <- c(0.1, 0.3, 0.5, 0.7, 0.9)

femalePercentiles <- quantile(female, percentiles)
malePercentiles <- quantile(male, percentiles)

#Create data frame with both 
percentileTable <- data.frame(
  Percentile = c("10th", "30th", "50th", "70th", "90th"), 
  Female = femalePercentiles,
  Male = malePercentiles
)
print(percentileTable)
```
## Question 3
Study the following boxplots showing population sizes by county. Which continent has the country with the biggest population size?

Asia has the country with the biggest population size because it has the tallest upper whisker with the highest point on the y-axis (next to 1000 million), which represents the max population size of the countries in each continent. 

## Question 4
Which continent has the largest median population size? 
Africa has the largest median population size because its median line is higher on the y-axis than the othercontinents. Median is represented by the horizontal line inside each box. 

## Quesion 5
What is median population size for African to the nearest million? 
Africa's median population size is about 10 million. 

## Question 6 
What proportion of countries in Europe have populations below 14 million? 
b. 0.75. 
This is because Europe's median life is below 14 million and the top of the box (75th percentile) is also slightly below 14 million. 

## Question 7 
If we use a log transformation, which continent shown above has the largest interquartile range?
Asia has the largest interquartile range because it has the tallest box. The box has the widest vertical range. 

## Question 8 
Load the height data and create a vector x with just the male heights. 
```{r}
library(dslabs)
x <- heights$height[heights$sex=="Male"]
```

What proportion of the data is between 69 and 72 inches (taller than 69, but shorter or equal to 72)? 
Hint: use a logical operator and mean.
```{r}
mean(x > 69 & x <= 72)
```
The mean is 0.3337438. 

